---
title: "Chapter 01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This chapter corresponds to "Introruction to Python" in the original book.


## Data types

We will use `class` function to know the variable types.
Notice that type of `10` is `numeric` not `integer`.
Append `L` to make it treated as an integer. 

```{r}
class(10)
class(10L)
class(2.718)
class("hello")
```

## Variables 

Like Python, R also supports dynamic typing.

```{r}
x <- 100L
y <- 3.14
x*y
class(x*y)
```

## List and vectors

For basic types such as numeric and character, we can make a vector by `c` function.
R uses one-base indexing (index starts with 1, not 0).

```{r}
a <- c(1,2,3,4,5)
a
length(a)
a[1]
a[5]
a[5] <- 99
a
```

What is more similar to python's list would be `list`, since it can have different
types of elements.
Use `[[` to access list elements.
```{r}
a <- list(1,2,3,4,5)
a
length(a)
a[[1]]
a[[5]]
a[[5]] <- 99
a
```


## Dictionary, and hash table

Vectors of lists can have names, with which we can use them like a dictionary.

```{r}
## Vector 
me <- c(height = 180)
me["height"]
me["weight"] <- 70
me

## List 
me <- list(height = 180)
me[["height"]]
me[["weight"]] <- 70
me
```

Alternatively, we can use `hash` function from `hash` library.

```{r}
library(hash)
me <- hash(height = 180)
me$height
me$weight <- 70
me
```



## Boolean or logicals

In R, boolean variables called `logical` and have values `TRUE` or `FALSE`.

```{r}
hungry <- TRUE
sleepy <- FALSE
class(hungry)
!hungry
hungry & sleepy
hungry | sleepy
```



## If-statement

```{r}
hungry <- TRUE
if (hungry) print("I'm hungry")
hungry <- FALSE
if (hungry) {
  print("I'm hungry")
} else {
  print("I'm not hungry")
  print("I'm sleepy")
}
```


## For loop

```{r}
for (i in 1:3) print(i)
```



## Functions

```{r}
hello <- function()
{
  print("Hello World!")
}
hello()

hello <- function(object)
{
  print(sprintf("Hello %s!", object))
}
hello("cat")
```



## Script File

To run a script file from R session, use `source` function.

```{r}
source("hungry.R")
```

To run a script file from console, call `Rscript`.

```bash
$ Rscript hungry.R 
[1] "I'm hungry!"
```


## Class

R has various ways to define classes.   In this exercise, we will use [`R6`](https://cran.r-project.org/web/packages/R6/vignettes/Introduction.html) implementation.
See more about object oriented programming in R, see [Hadley's Advanced R](http://adv-r.had.co.nz/OO-essentials.html).

```{r}
library(R6)
Man <- R6Class(
  "Man",
  public = list(
    name = "",
    initialize = function(name) { self$name <- name },
    hello = function() { print(sprintf("Hello %s!", self$name))  },
    goodbye = function()  { print(sprintf("Gooe-bye %s!", self$name)) }
  )
)

m <- Man$new("David")
m$hello()
m$goodbye()
```


## Matrix and array

What's similar to numpy's arrays in R is `vector` (1d), `matrix` (2d) and `array` (3d+).

```{r}
x <- c(1,2,3)
y <- c(2,4,6)
x+y
x*y
x/y
x/2
```

```{r}
A <- matrix(c(1,2,3,4), nrow=2, ncol=2, byrow=TRUE)
A
dim(A)

B <- matrix(c(3,0,0,6), nrow=2, ncol=2, byrow=TRUE)
A+B
A*B
A*10
```


## Broadcast

R's broadcast rule is the opposite of that of python.
```{r}
A <- matrix(c(1,2,3,4), nrow=2, ncol=2, byrow=TRUE)
B <- c(10,20)
A*B

# to reproduce the book's result,
A <- matrix(c(1,2,3,4), nrow=2, ncol=2, byrow=FALSE)
B <- c(10,20)
t(A*B)
```


## Array

```{r}
x <- array(1, dim=c(4,3,2))
x
y <- 1:8
x*y
```


## Access matrix/array elements

```{r}
X <- matrix(c(51,55,14,19,0,4), nrow=3, ncol=2, byrow=TRUE)
X
X[1,]
X[1,2]
for (i in 1:nrow(X)) print(X[i,])
```

Elements are tracked by "column" first way in R.
So flattening behavior is different from Python.
```{r}
dim(X) <- prod(dim(X))
X

# to reproduce the book result,
X <- matrix(c(51,55,14,19,0,4), nrow=3, ncol=2, byrow=FALSE)
dim(X) <- prod(dim(X))
X
X[c(1,3,5)]
X > 15
X[X > 15]
```


## Plot

R has various ways to visualize data.  In this exercise, we will use `ggplot2` library.
```{r}
library(ggplot2)
x <- seq(0, 6, 0.1)
y <- sin(x)
qplot(x, y, geom="line")

y1 <- sin(x)
y2 <- cos(x)
tmp <- rbind(data.frame(x=x, y=y1, func="sin", stringsAsFactors=FALSE),
             data.frame(x=x, y=y2, func="cos", stringsAsFactors=FALSE))
ggplot(tmp, aes(x, y, linetype=func)) + geom_line() + ggtitle("Sin & Cos")
```


## Image plot

To read a png format image file, use `readPNG` from `png` library.
It is loaded as a 3d array of shape `(nrow, ncol, channel)`.
To visualize image, we can use `grid.raster` from `grid` library.
```{r}
library(png)
img <- readPNG("dataset/lena.png")
str(img)

library(grid)
grid.raster(img)
```

`grid.raster` does not return graphic object.  To use image plot as if it is a `ggplot` objects, we will use the following custom function.

```{r}
image_plot <- function(arr)
{
  out <- grid::rasterGrob(arr)

  out <- ggplot2::qplot(0.5, 0.5, xlim=c(0,1), ylim=c(0,1)) +
    ggplot2::theme_void() + ggplot2::xlab('') + ggplot2::ylab('') +
    ggplot2::annotation_custom(out)

  out
}
image_plot(img)
class(image_plot(img))
```
